---
title: 'Learning R HW 1'
author: 'Gregory Schlett'
date: '02-07-2025'
output: html_document
---

```{r}
plot(typ, main = "Employee Response")
```


```{r}
install.packages(fpp)
library(fpp)
typ <- Typical_Employee_Survey_Data
typ
```

library(readxl)
Typical_Employee_Survey_Data <- read_excel("Downloads/Typical_Employee_Survey_Data.xlsx")

names(Typical_Employee_Survey_Data) = c('Age','Gender','Satisfaction','Characteristic','YearsWithComp','PromotionLikelihood','DecisionMaking','BudgetDecisions','CompanyPride','JobOfferScenario','Relationships')
Typical_Employee_Survey_Data

boxplot(Typical_Employee_Survey_Data$Satisfaction, Typical_Employee_Survey_Data$Characteristic, Typical_Employee_Survey_Data$YearsWithComp, Typical_Employee_Survey_Data$CompanyPride, Typical_Employee_Survey_Data$JobOfferScenario)

Typical_Employee_Survey_Data$Gender <- as.factor(Typical_Employee_Survey_Data$Gender)
class(Typical_Employee_Survey_Data$Gender)

print(is.factor(Typical_Employee_Survey_Data$Gender))

Typical_Employee_Survey_Data$PromotionLikelihood <- as.factor(Typical_Employee_Survey_Data$PromotionLikelihood)
class(Typical_Employee_Survey_Data$PromotionLikelihood)

# Quicker way to transform columns into factors
factor_cols <- c("Gender", "Satisfaction", "Characteristic", "PromotionLikelihood", "DecisionMaking", "BudgetDecisions", "CompanyPride", "JobOfferScenario", "Relationships")
data[factor_cols] <- lapply(data[factor_cols], as.factor)

class(Typical_Employee_Survey_Data)

stem(Typical_Employee_Survey_Data$Age)
stem(Typical_Employee_Survey_Data$Satisfaction)
stem(Typical_Employee_Survey_Data$Characteristic)
stem(Typical_Employee_Survey_Data$YearsWithComp)
stem(Typical_Employee_Survey_Data$CompanyPride)
stem(Typical_Employee_Survey_Data$JobOfferScenario)
stem(Typical_Employee_Survey_Data$Relationships)

#Age Plot
barplot(Typical_Employee_Survey_Data$Age, main = "Age", xlab = "Employee", ylab = "Age (in years)", col = "Green", names.arg = Typical_Employee_Survey_Data$Number)
Typical_Employee_Survey_Data$Age <- as.numeric(Typical_Employee_Survey_Data$Age)
barplot(Typical_Employee_Survey_Data$Age, main = "Age Distribution", xlab = "Employee", ylab = "Age (in years)", col = "green", names.arg = Typical_Employee_Survey_Data$Age, las = 2)

#Gender Chart
Typical_Employee_Survey_Data$Gender <- as.factor(Typical_Employee_Survey_Data$Gender)
gender_count <- table(Typical_Employee_Survey_Data$Gender)

pie(gender_count, labels = names(gender_count), main = "Gender", col = rainbow(length(gender_count)))

print(gender_count)
names(gender_count) <- c('Male', 'Female')
print(gender_count)

pie(gender_count, labels = names(gender_count), main = "Gender", col = rainbow(length(gender_count)))

#Satisfaction 
barplot(Typical_Employee_Survey_Data$Satisfaction, main = "Employee Satisfaction", xlab = "Employee", ylab = "Satisfaction Response", col = "Green", names.arg = Typical_Employee_Survey_Data$Number)

str(Typical_Employee_Survey_Data)

names(Typical_Employee_Survey_Data)

summary(Typical_Employee_Survey_Data)
boxplot.stats(Typical_Employee_Survey_Data$Age)

x1 = Typical_Employee_Survey_Data$Age
y1 = Typical_Employee_Survey_Data$Gender

plot(Typical_Employee_Survey_Data$Age)
plot(x1, main="Age and Gender", xlab="Age", ylab="Gender", col="red", cex=2)
points(y1, col="blue", cex=2)

plot(Typical_Employee_Survey_Data$CompanyPride)
plot(Typical_Employee_Survey_Data$Satisfaction)
plot(Typical_Employee_Survey_Data$JobOfferScenario)





